comgen-outline

### 2017-06-08 paper/generics-paper.Rmd
fix citations

---
title             : "Communicating generalizations about categories, events, and causes"
shorttitle        : "Communicating generalizations"

@@

Bayesian cognitive models are used to represent flexible, structured knowledge of the world and formalize the inferences that people draw using such conceptual representations [e.g., @Kemp2008; @Goodman2015Chapter].
The basic currency of inference in these models is probability.
In the case of generic language, we posit that conceptual factors may influence the meaning of a generic statement via its influence on subjective probability.
If that's true, a simple, semantic theory (e.g., a threshold on subjective probability) may be enough to formalize the core meaning of a generic statement $u_{gen}$:

@@

```{r simulations, fig.width = 8, fig.height=5}
all.simulations %>%
  mutate(Distribution = paste(Distribution, utterance, sep = "_")) %>%
  filter(Distribution != "Some", Distribution != "Most", Distribution != "Prior_generic") %>%
  mutate(Distribution = factor(Distribution, levels = 
                             c("Prior_some",  "Posterior_Generalization_generic", "Posterior_Generalization_some"),
                                labels = c("Prior/Silence",  "Generalization (Uncertain threshold)", "Fixed threshold")),
         PriorShape = factor(PriorShape, levels = c("uniform", "barks","laysEggs","carriesMalaria","isFemale"),
                             labels = c("uniform", "bark","lay eggs","carry malaria","are female")
                                  #"eatsFood","areSick", "uniformRare", 
                                  #"carriesMalaria"
                                  )
         ) %>%
  ggplot(., aes(x = value, #y = prob, 
                #lty = Distribution, 
                color = Distribution, fill = Distribution))+
    geom_density(#stat= 'identity', 
                aes(y = ..scaled..),
                 size = 0.6, alpha = 0.7, adjust = 1.5)+
    #geom_bar(stat = 'identity', position = position_dodge(), color = 'black')+
    theme_few() +
    #scale_color_solarized()+
    #scale_fill_solarized()+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
   scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_linetype_manual(values = c(3, 4, 1))+
    scale_x_continuous(breaks = c(0, 0.5, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Degree of probability h") +
    ylab("Scaled probability density") +
  facet_grid(PriorShape~Distribution, scales = 'free')+
  guides(fill = F, color = F)+
  theme(strip.text.y = element_text(angle = 0))
```

@@

Generalizations about categories (i.e., *generic language*) have received substantial interest in psychology for their role in concept formation [@Gelman2004] and stereotype propogation [@Rhodes2012].
Generic language (e.g., *Birds fly.*) exhibits substantial flexibility in meaning and truth conditions as described above.
<!-- At first glance, generics feel like universally-quantified statements as in *All* birds fly.  -->
<!-- Unlike universals, however, generics are resilient to counter-examples (e.g., *Birds fly* seems true even in the face of flightless birds like kiwis or penguins). -->
<!-- *Most* birds fly seems more reasonable, and yet most mosquitos *do not* carry malaria while *Mosquitos carry malaria* seems like a true statement. -->
<!-- Finally, the probability of a bird being female is much higher than the probability of a mosquito carrying malaria, but *Birds are female* seems weird to say. -->
<!-- *Birds lay eggs*, however, is fine, even though the rate is the same as *are female* (only females lay eggs). -->

@@

As a first test of our theory, we wish to explain the flexibility in endorsements of generics in terms of the relevant prior knowledge about properties and show how it systematically predicts endorsements of familiar generic statements (Expt. 1).
In Experiment 3, we replicate and extend a paradigm used by @Cimpian2010 to demonstrate the asymmetry between endorsements and interpretations and find that a similar pattern is predicted by our model.

@@

To compare the model to empirical truth judgments, we first measure the prior distribution over prevalence of these properties (Expt. 1a).
We hypothesize that this prior distribution encodes probabilities associated with deeper conceptual knowledge [e.g., of the kind discussed by @Prasada2013].
This would be one way in which our theory touches upon other psychological accounts of generic language [cf., @Prasada; @Gelman; @Leslie].
If, indeed, there are such things as different types of kind--property relations (e.g., *principled* v. *accidental relations*), we would expect the prevalence prior data to follow a multi-modal, mixture distribution where different modes or components of the prior are the results of the different types of relations.
As a simple of this hypothesis, we compare a structured (i.e., mixture) model of the prior data to an unstructured one.

@@

generics.endorsement.models <- bind_rows(
  d.gen.endorse.regression %>% 
  mutate(prediction_lower = prediction,
         prediction_upper = prediction),
  left_join(
    m.gen.endorse.rsa %>% 
    rename(prediction = MAP, prediction_lower = cred_lower, prediction_upper = cred_upper),
    d.gen.endorse.regression %>% 
    select(sentence, Property, Category, prevalence, cuevalidity)
  )
) %>%
  mutate(src = factor(src, levels = c( 
                                      "generics_model",
                                      "some_model",
                                      "regression_Prev", 
                                      "regression_Prev_Cuevalidity"
                                     ),
                      labels = c(
                                 "Speaker model - uncertain semantics",
                                 "Speaker model - fixed semantics",
                                 "Prevalence alone", "Prevalence & Cue validity"
                                 ))) %>%
  ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = prevalence ))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human generic endorsement")+
  facet_wrap(~src, nrow = 2)+
  theme(legend.position = "bottom")

@@

```{r generics-endorsement-figure, fig.width=7, fig.height=5}
grid.arrange(generics.endorsement.spectrum, 
             generics.endorsement.models,
             generics.endorsement.insets, ncol = 2,
             layout_matrix = cbind(c(1,1,3), c(2,2,3)))
```

@@

Learning that an event consistently tends to result in another event serves as a foundation for an intuitive theory of causality (CITE noah's and kemp's causality and causal models papers?).
Communicating generalizations about causally related events is potentially highly useful because no one individual can run all the experiments that she wants to run in a life time. 
The history of civilizations has shown that it is incredibly useful to divide labor: Some members can investigate modifications to the food preparation procedures while others can examine different ways of cutting stone.

Our framework for communicating generalizations makes predictions about statements conveying genericty about a category of causal events. 
In this domain, an instance of the category is a single causal event (e.g., adding a particular amount of a particular kind of yeast to an actual piece of dough at a particular time of the day). 
@Gerstenberg
The study of *elemental causal induction* is itself very subtle, depending on an observer's background theory of causality. [@Griffiths2005; @Schulz].

@@

The theory we present here has a number of desirable features that go beyond alternative accounts.
The first is that our theory is accompanied by a formal model, which makes precise quantitative predictions about endorsments of statements that convey generalizations.
The formal Bayesian model provides a clean separation of the semantics of a generalization from background, world knowledge.
This is critical to understanding the contribution of a generalization in language above and beyond background knowledge.
We believe this has been a stumbling block for formal semantic models that try to argue in one or another way for a statistical or probabilty based approach [@Pelletier; @Cohen1999; @Nickel].
As @Nickel2013 notes, genericity is particularly difficult to study because it intersects with many other phenomena in language (e.g., gradable adjectives in *Giraffes are tall*).
We argue that it also intersects with world knowledge in subtle ways, and it is critical to accurately measure and describe world knowledge before making claims about the implications of a generalization *per se*.

@@

This may be throwing out the baby with the bathwater.
When we consider the acquisition problem, there are two aspects of meaning that a language learner must acquire for the semantics of a quantified utterance: (1) that the meaning is a threshold-function (i.e., $\denote{u}(p, \theta) = \{p :p > \theta\}$ and (2) the fixed value of the threshold (e.g., $\theta = 0$ for "some", $\theta = 0.5$ for "most").
If there is no fixed value of the threshold, then a uniform prior over the threshold in the truth-functional threshold semantics is mathematically equivalent to a *soft semantics* wherein the degree to which the utterance is true is *the degree itself*, normalized by the prior distribution over the degree (i.e., the comparison class).

$$
\int_{0}^{1} \delta_{p > \theta} \diff \theta =  p
$$

@@

Finally, (4) is used to argue that even if the property is widespread (perhaps in a soft sense, as in our model, relative to other kinds or groups), we might find it avervise to actually assert the generic.
That is, even if it were the case that, because of a variety socio-cultural reasons, women tend to be submissive, we might still not want to say *Women are submissive* [@Haslanger]. 
In our Model Simulations section, we showed that the generic often can imply the property is widespread, even with uninformed prior beliefs about the property. 
If the generic conveys the property is in fact more widespread than the speaker herself might believe, this can lead to a distortion of the truth [cf., @Cimpian2010].
The current formulation of the model is relatively ambivalent the deviations of the listener's beliefs from the speaker's beliefa.
Other models in the RSA framework have identified states of the world with different subjective values (e.g., how good of a job a listener did on a task), and speakers with utilities that extend beyond conveying information can produce qualitatively different patterns of utterances [e.g., polite speech; @Yoon2016; @Yoon2017].
If our model believed that states of the world where the prevalence of the feature is high (e.g., a world where many women were submissive) were undesirable, then a speaker who considered both the informational content of the utterance and the utility in the states of the world implied by that utterance would too find it reprehensible to make such generic claims.

>>>>

---
title             : "Communicating generalizations"
shorttitle        : "Communicating generalizations"

@@

Bayesian cognitive models are used to represent flexible, structured knowledge of the world and formalize the inferences that people draw using such conceptual representations [e.g., @Kemp2008; @Goodmanconcepts].
The basic currency of inference in these models is probability.
In the case of generic language, we posit that conceptual factors may influence the meaning of a generic statement via its influence on subjective probability.
If that's true, a simple, semantic theory (e.g., a threshold on subjective probability) may be enough to formalize the core meaning of a generic statement $u_{gen}$:

@@

```{r simulations, fig.width = 9, fig.height=5}
all.simulations %>%
  mutate(Distribution = paste(Distribution, utterance, sep = "_")) %>%
  filter(Distribution != "Some", Distribution != "Most", Distribution != "Prior_generic") %>%
  mutate(Distribution = factor(Distribution, levels = 
                             c("Prior_some",  "Posterior_Generalization_generic", "Posterior_Generalization_some"),
                                labels = c("Prior/Silence",  "Generalization (Uncertain threshold)", "Fixed threshold")),
         PriorShape = factor(PriorShape, levels = c("uniform", "barks","laysEggs","carriesMalaria","isFemale"),
                             labels = c("uniform", "bark","lay eggs","carry malaria","are female")
                                  #"eatsFood","areSick", "uniformRare", 
                                  #"carriesMalaria"
                                  )
         ) %>%
  ggplot(., aes(x = value, #y = prob, 
                #lty = Distribution, 
                color = Distribution, fill = Distribution))+
    geom_density(#stat= 'identity', 
                aes(y = ..scaled..),
                 size = 0.6, alpha = 0.7, adjust = 1.5)+
    #geom_bar(stat = 'identity', position = position_dodge(), color = 'black')+
    theme_few() +
    #scale_color_solarized()+
    #scale_fill_solarized()+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
   scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_linetype_manual(values = c(3, 4, 1))+
    scale_x_continuous(breaks = c(0, 0.5, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Degree of probability h") +
    ylab("Scaled probability density") +
  facet_grid(PriorShape~Distribution, scales = 'free')+
  guides(fill = F, color = F)+
  theme(strip.text.y = element_text(angle = 0))
```

@@

Generalizations about categories (i.e., *generic language*) have received substantial interest in psychology for their role in concept formation [@Gelman2004] and stereotype propogation [@Rhodes2012].
Generic language (e.g., *Birds fly.*) exhibits substantial flexibility in meaning and truth conditions as described above.
As a first test of our theory, we wish to explain the flexibility in endorsements of generics (e.g., "Birds lay eggs" vs. "Birds are female"; "Mosquitos carry malaria") in terms of the relevant prior knowledge about properties.
We do this by measuring the prior belief distribution over the relevant probabilities $P(h)$ used in Eq. \ref{eq:L0} as well as the target probabilitiy that the speaker model seeks to convey $h$ in Eq. \ref{eq:S1}.
We show how these measurements systematically predict, via our model, endorsements of generic statements about familiar categories (Expt. 1b).

<!-- At first glance, generics feel like universally-quantified statements as in *All* birds fly.  -->
<!-- Unlike universals, however, generics are resilient to counter-examples (e.g., *Birds fly* seems true even in the face of flightless birds like kiwis or penguins). -->
<!-- *Most* birds fly seems more reasonable, and yet most mosquitos *do not* carry malaria while *Mosquitos carry malaria* seems like a true statement. -->
<!-- Finally, the probability of a bird being female is much higher than the probability of a mosquito carrying malaria, but *Birds are female* seems weird to say. -->
<!-- *Birds lay eggs*, however, is fine, even though the rate is the same as *are female* (only females lay eggs). -->

@@

@@

To compare the model to empirical truth judgments, we first measure the prior distribution over prevalence of these properties (Expt. 1a).
We hypothesize that this prior distribution encodes probabilities associated with deeper conceptual knowledge [e.g., of the kind discussed by @Prasada2013].
This would be one way in which our theory touches upon other psychological accounts of generic language [cf., @Prasada2013; @Gelman2003; @Leslie2007].
If, indeed, there are such things as different types of kind--property relations (e.g., *principled* v. *accidental relations*), we would expect the prevalence prior data to follow a multi-modal, mixture distribution where different modes or components of the prior are the results of the different types of relations.
As a simple test of this hypothesis, we compare a structured (i.e., mixture) model of the prior data to an unstructured one.

@@

generics.endorsement.models <- bind_rows(
  d.gen.endorse.regression %>% 
  mutate(prediction_lower = prediction,
         prediction_upper = prediction),
  left_join(
    m.gen.endorse.rsa %>% 
    rename(prediction = MAP, prediction_lower = cred_lower, prediction_upper = cred_upper),
    d.gen.endorse.regression %>% 
    select(sentence, Property, Category, prevalence, cuevalidity)
  )
) %>%
  mutate(src = factor(src, levels = c( 
                                      "generics_model",
                                      "some_model",
                                      "regression_Prev", 
                                      "regression_Prev_Cuevalidity"
                                     ),
                      labels = c(
                                 "Uncertain semantics model",
                                 "Fixed semantics model",
                                 "Prevalence alone", "Prevalence + Cue validity"
                                 ))) %>%
  ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = prevalence ))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human generic endorsement")+
  facet_wrap(~src, nrow = 2)+
  theme(legend.position = "bottom")

@@

```{r generics-endorsement-figure, fig.width=11, fig.height=8}
grid.arrange(generics.endorsement.spectrum, 
             generics.endorsement.models,
             generics.endorsement.insets, ncol = 2,
             layout_matrix = cbind(c(1,1,3), c(2,2,3)))
```

@@

Learning that an event consistently tends to result in another event serves as a foundation for an intuitive theory of causality [@Goodman2011; @Kemp2010a].
Communicating generalizations about causally related events is potentially highly useful because no one individual can run all the experiments that she wants to run in a life time. 
The history of civilizations has shown that it is incredibly useful to divide labor: Some members can investigate modifications to the food preparation procedures while others can examine different ways of cutting stone.

Our framework for communicating generalizations makes predictions about statements conveying genericty about a category of causal events. 
In this domain, an instance of the category is a single causal event (e.g., adding a particular amount of a particular kind of yeast to an actual piece of dough at a particular time of the day). 
[Gerstenberg]
The study of *elemental causal induction* is itself very subtle, depending on an observer's background theory of causality. [@Griffiths2005; @Schulz2008a].

@@

The theory we present here has a number of desirable features that go beyond alternative accounts.
The first is that our theory is accompanied by a formal model, which makes precise quantitative predictions about endorsments of statements that convey generalizations.
The formal Bayesian model provides a clean separation of the semantics of a generalization from background, world knowledge.
This is critical to understanding the contribution of a generalization in language above and beyond background knowledge.
We believe this has been a stumbling block for formal semantic models that try to argue in one or another way for a statistical or probabilty based approach [@Cohen1999; @Nickel2008].
As @Nickel2016 notes, genericity is particularly difficult to study because it intersects with many other phenomena in language (e.g., gradable adjectives in *Giraffes are tall*).
We argue that it also intersects with world knowledge in subtle ways, and it is critical to accurately measure and describe world knowledge before making claims about the implications of a generalization *per se*.

@@

This may be throwing out the baby with the bathwater.
When we consider the acquisition problem, there are two aspects of meaning that a language learner must acquire for the semantics of a quantified utterance: (1) that the meaning is a threshold-function (i.e., $\denote{u}(h, \theta) = \{h :h > \theta\}$ and (2) the fixed value of the threshold (e.g., $\theta = 0$ for "some", $\theta = 0.5$ for "most").
If there is no fixed value of the threshold, then a uniform prior over the threshold in the truth-functional threshold semantics is mathematically equivalent to a *soft semantics* wherein the degree to which the utterance is true is *the degree itself*, normalized by the prior distribution over the degree (i.e., the comparison class).

$$
\int_{0}^{1} \delta_{h > \theta} \diff \theta =  h
$$

@@

Finally, (4) is used to argue that even if the property is widespread (perhaps in a soft sense, as in our model, relative to other kinds or groups), we might find it avervise to actually assert the generic.
That is, even if it were the case that, because of a variety socio-cultural reasons, women tend to be submissive, we might still not want to say *Women are submissive* [@Haslanger2011]. 
In our Model Simulations section, we showed that the generic often can imply the property is widespread, even with uninformed prior beliefs about the property. 
If the generic conveys the property is in fact more widespread than the speaker herself might believe, this can lead to a distortion of the truth [cf., @Cimpian2010].
The current formulation of the model is relatively ambivalent the deviations of the listener's beliefs from the speaker's beliefa.
Other models in the RSA framework have identified states of the world with different subjective values (e.g., how good of a job a listener did on a task), and speakers with utilities that extend beyond conveying information can produce qualitatively different patterns of utterances [e.g., polite speech; @Yoon2016; @Yoon2017].
If our model believed that states of the world where the prevalence of the feature is high (e.g., a world where many women were submissive) were undesirable, then a speaker who considered both the informational content of the utterance and the utility in the states of the world implied by that utterance would too find it reprehensible to make such generic claims.
